\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-2\baselineskip}

### 2017-09-18 paper/ms.tex
more response to referee

We then sample the joint posterior probability (\eqalt{joint-post}) using
\emcee\ \citep{Goodman:2010, Foreman-Mackey:2013}.
We initialize \exampleinwalkers~walkers\footnote{\responseb{``Walker'' is the
name given by \citet{Goodman:2010} and \citet{Foreman-Mackey:2013} to each
parallel MCMC chain in the ensemble sampling algorithm used in \emcee.}} by
sampling from a three-dimensional isotropic Gaussian centered the MAP
parameter vector and with a standard deviation of $10^{-4}$ in each dimension.
We then run \exampleinburn~steps of burn-in and \exampleinsteps~steps of MCMC.
To assess convergence, we estimate the mean integrated autocorrelation time of
the chain across parameters \citep{Sokal:1989, Goodman:2010} and find that the
chain results in \exampleineff~effective samples.

@@

\begin{figure}[hptb]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/rotation/rotation.pdf}
\caption{Inferred constraints on a quasiperiodic GP model using the covariance
    function in \eq{rot-kernel} and two quarters of \kepler\ data.
(top left) The \kepler\ data (black points) and the MAP model
    prediction (blue curve) for a 60~day subset of the data used.
    The solid blue line shows the predictive mean and the blue contours show
    the predictive standard deviation.
(bottom left) The residuals between the mean predictive model and the data
    shown in the top left figure.
(right) The posterior constraint on the rotation period of KIC~1430163 using
    the dataset and model from \Figure{rotation}.
    The period is the parameter $P_\mathrm{rot}$ in \eq{rot-kernel} and this
    figure shows the posterior distribution marginalized over all other
    nuisance parameters in \eq{rot-kernel}.  This is consistent with the
    published rotation period made using the full
    \kepler\ baseline shown as a vertical gray line \citep{Mathur:2014}.
    \figurelabel{rotation}}
\end{center}
\end{figure}

@@

This model requires about 10~CPU minutes to run the MCMC to convergence.
This is more computationally intensive than traditional methods of measuring
asteroseismic oscillations, but it is much cheaper than the same analysis
using a general direct GP solver.
For comparison, we estimate that repeating this analysis using a general
Cholesky factorization implemented as part of a tuned linear algebra
library\footnote{We use the Intel Math Kernel Library
\url{https://software.intel.com/en-us/intel-mkl}} would require about 15~CPU
hours.
An in-depth discussion of the benefits of rigorous probabilistic inference of
asteroseismic parameters in the time domain is beyond the scope of this paper,
but we hope to revisit this opportunity in the future.
}

@@

\begin{figure}[!p]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/astero/astero-corner.pdf}
\caption{The probabilistic constraints on $\nu_\mathrm{max}$ and $\Delta \nu$
    from the inference shown in \Figure{astero} compared to the published
    value (error bar) based on several years of \kepler\ observations
    \citep{Pinsonneault:2014}.
    The two-dimensional contours show the 0.5-, 1-, 1.5, and 2-sigma credible
    regions in the marginalized planes and the histograms along the diagonal
    show the marginalized posterior for each parameter.
    \figurelabel{astero-corner}}
\end{center}
\end{figure}

@@

\response{
This example is different from all the previous examples~--~both simulated and
real~--~because, in this case, do not set the deterministic
mean function $\mu_\bvec{\theta}$ to zero.
Instead, we make inferences about $\mu_\bvec{\theta}$ because it is a
physically significant function and the parameters are fundamental properties
of the system.
This is an example using real data~--~because we use the light curve from
\sect{rotation}~--~but we multiply these data by a simulated transiting
exoplanet model with known parameters.
This allows us to show that we can recover the true parameters of the planet
even when the transit signal is superimposed on the real variability of a
star.
}
GP modeling has been used extensively for this purpose throughout the
exoplanet literature \citep[for example][]{Dawson:2014, Barclay:2015,
Evans:2015, Foreman-Mackey:2016b, Grunblatt:2016}.

@@

The continuous autoregressive moving average (CARMA) models introduced to the
astrophysics literature by \citet{Kelly:2014} share many features with
\celerite.
CARMA models are restricted to one-dimensional problems and the likelihood
function for a CARMA model can be solved in $\mathcal{O}(N\,J^2)$ using a
Kalman filter \citep{Kelly:2014}.
The kernel function for a CARMA$(J,\,K)$ model is
\begin{eqnarray}\eqlabel{carma-kernel}
k_\mathrm{CARMA}(\tau) = \sum_{j=1}^J A_j\,\exp\,(r_j\,\tau)
\end{eqnarray}
where
\begin{eqnarray}\eqlabel{carma-coeff}
A_j = \sigma^2 \,\frac{\left[\sum_{k=0}^K\beta_k\,{(r_j)}^k\right]\,
    \left[\sum_{k=0}^K\beta_k\,{(-r_j)}^k\right]}
    {-2\,\mathrm{Re}(r_j)\,\prod_{k=1,\,k \ne j}^{J}(r_k-r_j)({r_k}^*+r_j)}
\end{eqnarray}
and $\sigma$, $\{r_j\}_{j=1}^J$ and $\{\beta_k\}_{k=1}^K$ are parameters of
the model; $\sigma$ is real, while the others are complex.
Comparing \eq{carma-kernel} to \eq{celerite-kernel-complex}, we can see that
every CARMA model corresponds to an equivalent \celeriteterm\ model and the
parameters $a_j$, $b_j$, $c_j$, and $d_j$ can be easily computed analytically.
The inverse statement is not as simple.
In practice, this means that \celerite\ could be trivially used to compute any
CARMA model.
Using the CARMA solver to compute a \celeriteterm\ model, however, requires
solving \eq{carma-coeff} numerically for a given set of $\{A_j\}_{j=1}^J$.

>>>>

We then sample the joint posterior probability (\eqalt{joint-post}) using
\emcee\ \citep{Goodman:2010, Foreman-Mackey:2013}.
We initialize \exampleinwalkers~walkers\footnote{\responseb{``Walker'' is the
name given by \citet{Goodman:2010} and \citet{Foreman-Mackey:2013} to each
parallel MCMC chain in the ensemble sampling algorithm used in \emcee.}} by
sampling from a three-dimensional isotropic Gaussian centered the MAP
parameter vector and with a standard deviation of $10^{-4}$ in each dimension.
We then run \exampleinburn~steps of burn-in and \exampleinsteps~steps of MCMC.
To assess convergence, we estimate the integrated autocorrelation time of the
chain \responseb{averaged across parameters} \citep{Sokal:1989, Goodman:2010}
and find that the chain results in \exampleineff~effective samples.

@@

\begin{figure}[hptb]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/rotation/rotation.pdf}
\caption{Inferred constraints on a quasiperiodic GP model using the covariance
    function in \eq{rot-kernel} and two quarters of \kepler\ data.
(top left) The \kepler\ data (black points) and the MAP model
    prediction (blue curve) for a 60~day subset of the data used.
    The solid blue line shows the predictive mean and the blue contours show
    the predictive standard deviation.
(bottom left) The residuals between the mean predictive model and the data
    shown in the top left figure.
(right) The posterior constraint on the rotation period of KIC~1430163 using
    the dataset and model from \Figure{rotation}.
    The period is the parameter $P_\mathrm{rot}$ in \eq{rot-kernel} and this
    figure shows the posterior distribution marginalized over all other
    nuisance parameters in \eq{rot-kernel}.
    \responseb{The 1-$\sigma$ error bar on this measurement is indicated with
    vertical dashed lines.}
    This result is consistent with the
    published rotation period made using the full
    \kepler\ baseline shown as a vertical gray line \citep{Mathur:2014}.
    \figurelabel{rotation}}
\end{center}
\end{figure}

@@

\responseb{For this example, about 10~CPU minutes were required to run the
MCMC.}
This is more computationally intensive than traditional methods of measuring
asteroseismic oscillations, but it is much cheaper than the same analysis
using a general direct GP solver.
For comparison, we estimate that repeating this analysis using a general
Cholesky factorization implemented as part of a tuned linear algebra
library\footnote{We use the Intel Math Kernel Library
\url{https://software.intel.com/en-us/intel-mkl}} would require about 15~CPU
hours.
An in-depth discussion of the benefits of rigorous probabilistic inference of
asteroseismic parameters in the time domain is beyond the scope of this paper,
but we hope to revisit this opportunity in the future.
}

@@

\begin{figure}[!p]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/astero/astero-corner.pdf}
\caption{The probabilistic constraints on $\nu_\mathrm{max}$ and $\Delta \nu$
    from the inference shown in \Figure{astero} compared to the published
    value (\responseb{blue error bars}) based on several years of \kepler\
    observations \citep{Pinsonneault:2014}.
    The two-dimensional contours show the 0.5-, 1-, 1.5, and 2-sigma credible
    regions in the marginalized planes and the histograms along the diagonal
    show the marginalized posterior for each parameter.
    \figurelabel{astero-corner}}
\end{center}
\end{figure}

@@

\response{
This example is different from all the previous examples~--~both simulated and
real~--~because, in this case, do not set the mean function $\mu_\bvec{\theta}$
to zero.
Instead, we make inferences about $\mu_\bvec{\theta}$ because it is a
physically significant function and the parameters are fundamental properties
of the system.
This is an example using real data~--~because we use the light curve from
\sect{rotation}~--~but we multiply these data by a simulated transiting
exoplanet model with known parameters.
This allows us to show that we can recover the true parameters of the planet
even when the transit signal is superimposed on the real variability of a
star.
}
GP modeling has been used extensively for this purpose throughout the
exoplanet literature \citep[for example][]{Dawson:2014, Barclay:2015,
Evans:2015, Foreman-Mackey:2016b, Grunblatt:2016}.

@@

The continuous autoregressive moving average (CARMA) models introduced to the
astrophysics literature by \citet{Kelly:2014} share many features with
\celerite.
CARMA models are restricted to one-dimensional problems and the likelihood
function for a CARMA model can be solved in $\mathcal{O}(N\,J^2)$ using a
Kalman filter \citep{Kelly:2014}.
The kernel function for a CARMA$(J,\,K)$ model is
\begin{eqnarray}\eqlabel{carma-kernel}
k_\mathrm{CARMA}(\tau) = \sum_{j=1}^{2\,J} A_j\,\exp\,(r_j\,\tau)
\end{eqnarray}
where
\begin{eqnarray}\eqlabel{carma-coeff}
A_j = \sigma^2 \,\frac{\left[\sum_{k=0}^K\beta_k\,{(r_j)}^k\right]\,
    \left[\sum_{k=0}^K\beta_k\,{(-r_j)}^k\right]}
    {-2\,\mathrm{Re}(r_j)\,\prod_{k=1,\,k \ne j}^{J}(r_k-r_j)({r_k}^*+r_j)}
\end{eqnarray}
and $\sigma$, $\{r_j\}_{j=1}^{J}$ and $\{\beta_k\}_{k=1}^K$ are parameters
of the model; $\sigma$ is real, while the others are complex.
Comparing \eq{carma-kernel} to \eq{celerite-kernel-complex}, we can see that
every CARMA model corresponds to an equivalent \celeriteterm\ model and the
parameters $a_j$, $b_j$, $c_j$, and $d_j$ can be easily computed
analytically.\footnote{\responseb{Each CARMA term with complex parameters
$A_j$ and $r_j$ corresponds to a \celeriteterm\ term with parameters $a_j =
2\,\mathrm{Re}(A_j)$, $b_j = 2\,\mathrm{Im}(A_j)$, $c_j = -\mathrm{Re}(r_j)$,
and $d_j = -\mathrm{Im}(r_j)$.}}
The inverse statement is not as simple.
In practice, this means that \celerite\ could be trivially used to compute any
CARMA model.
Using the CARMA solver to compute a \celeriteterm\ model, however, requires
solving \eq{carma-coeff} numerically for a given set of $\{A_j\}_{j=1}^J$.
